Spatial variation in predator diel activity patterns -- feral cats avoid red foxes in time, not space

### 2022-04-14 manuscript/main_text.tex
post-phd-getting-ready-for-submission

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1.1in]{geometry}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Spatial variation in predator diel activity patterns -- feral cats avoid red foxes in time, not space},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

@@

We compiled data from multiple smaller-scale camera-trap studies across two regions in south-west Victoria, Australia: the Glenelg region and Otway Ranges (Fig. \ref{fig:diel-map}). Introduced foxes and cats are the only medium-large functional mammalian terrestrial carnivores here: native dingoes \emph{Canis familiaris} are long-absent throughout, while tiger quolls \emph{Dasyurus maculatus} are long-absent in the Glenelg region and likely functionally extinct in the Otway Ranges (last confirmed sighting in 2014). In broad sections of each region, government land managers conduct ongoing targeted lethal fox control for biodiversity conservation. Poison-baits containing 3 mg of sodium mono-fluroacetate (`1080') are buried at a depth of 12 - 15 cm at 1-km intervals along accessible forest tracks and roads. Different road densities result in variable densities of poison-baits. Managers also frequently implement prescribed fire across both regions, primarily to reduce fuel loads to prevent large wildfires.

@@

In the Glenelg region, large patches of natural vegetation are fragmented by pastoral farming and residential properties (Fig. \ref{fig:diel-map}). Foxes in three distinct forest blocks in this region have been subject to poison-baiting since October 2005, with fortnightly bait replacements (Robley \emph{et al.} 2014). These forest blocks, along with three similar, unbaited forest blocks to the north have been simultaneously surveyed annually under the `Glenelg Ark' fox control program since 2005 (40 sites per block; Robley \emph{et al.} 2020). Hair-tubes were used to monitor species from 2005 - 2013 (presented in Robley \emph{et al.} 2014), replaced by camera-traps from 2013; here we present camera-trap data from 2013 - 2019; Robley \emph{et al.} (2020){]}. We also included a further 425 camera-trap deployments at unique locations from early 2018 (M.W.R., PhD surveys). This totals 2039 camera-trap deployments in the Glenelg region, collected in a control-impact experimental design (foxes had been continuously controlled for at 8 - 14 years in the treatment landscapes at the time of these surveys).

@@

\hypertarget{camera-trap-set-ups}{%
\paragraph{Camera-trap set-ups}\label{camera-trap-set-ups}}

@@

\hypertarget{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-vegetation-types-model-2}{%
\paragraph{How does predator overall activity and diel activity patterns vary across vegetation types? (model 2)}\label{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-vegetation-types-model-2}}

@@

\hypertarget{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-space-model-1-1}{%
\subsection{How does predator overall activity and diel activity patterns vary across space? (model 1)}\label{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-space-model-1-1}}

@@

Variation in diel activity patterns across space, as well as differences in overall activity between the predators, was strongest in the Otway Ranges. For example, overall fox activity (Fig. \ref{fig:diel-space-o-fox}) and diel activity pattern strength(Fig. \ref{fig:diel-space}b) were lowest in the south-west Otway Ranges, while feral cat overall activity (Fig. \ref{fig:diel-space-o-cat}) and diel activity pattern strength (Fig. \ref{fig:diel-space}b) were highest in that subregion.

\hypertarget{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-vegetation-types-model-2-1}{%
\subsection{How does predator overall activity and diel activity patterns vary across vegetation types? (model 2)}\label{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-vegetation-types-model-2-1}}

@@

Diel activity patterns for foxes were similar across all vegetation type except wet forests; in wet forests, foxes were consistently active throughout the daily cycle (Fig. \ref{fig:diel-veg}a). On the other hand, cats were nocturnal (and most active) in wet forests, but largely crepuscular in all other vegetation types (Fig. \ref{fig:diel-veg}b). For both predators, the random effect for region (Glenelg or Otways) in the vegetation models shrank to near-zero=, indicating all variation between the regions was explained by the vegetation covariate and site random intercept.

\hypertarget{do-feral-cats-avoid-foxes-in-space-or-time-model-3-1}{%
\subsection{Do feral cats avoid foxes in space or time? (model 3)}\label{do-feral-cats-avoid-foxes-in-space-or-time-model-3-1}}

@@

\caption{Variation in mean feral cat \textit{Felis catus} activity (a) and associated uncertainty estimates (b) in response to count of ’independent’ red fox Vulpes vulpes detections (log-transformed and survey effort adjusted) across each ’habitat type’ in south-west Victoria, Australia (model 3). Grey vertical lines represent average sunrise and sunset times. In the Glenelg region, there were more feral cat detections where there were more fox detections, but cat peak diel activity shifted from crepuscular night to pre-dawn and midday (a). In the Otway Ranges, feral cat activity also peaked during the day where fox activity was high in dry vegetation types (b), but was more nocturnal where fox activity was high in the rainforests and wet forests (c).}\label{fig:diel-cat-fox}
\end{figure}

@@

A key question in ecological theory is whether animals are evolutionary hardwired to occupy particular temporal niches, or have circadian rhythms that are responsive to changing environmental conditions and interactions with other species (Schoener 1974; Daan 1981; Lima \& Dill 1990). Here we demonstrate that diel activity patterns are not fixed, but vary across space based on landscape context and fear. In our study, sympatric invasive predators had similar diel activity patterns when averaged across broad regions (i.e., high circular overlap; Fig. \ref{fig:diel-space-time-marginal}; as did Roshier \& Carter 2021), but behaviours varied considerably within landscapes. Fox daily activity patterns were most strongly tied to the daily cycle in dry habitat types, but showed little diel activity pattern in the wet forests (Fig. \ref{fig:diel-veg}a). In contrast, cats were mostly nocturnal in wet forests but crepuscular in dry vegetation types (Fig. \ref{fig:diel-veg}b). Within broad habitat types, cats altered their diel activity patterns at sites with higher fox activity (Fig. \ref{fig:diel-cat-fox}). Control programs that reducing invasive fox activity are therefore likely to change cat diel activity patterns, which in turn may alter cat impacts on native prey species. Quantifying changes in diel activity patterns provides important context for understanding species interactions, which is key for effective ecosystem management (Gaynor \emph{et al.} 2021).

Shifting diel activity patterns may facilitate spatial coexistence of dominant and subordinate species (Carothers \& Jaksić 1984). For cats, altering diel activity patterns to less-preferred times of day may be worthwhile to persist in high-quality habitat. Few studies have demonstrated predator-induced shifts in diel activity (Kronfeld-Schor \& Dayan 2003), but notably, ship rats \emph{Rattus norvegicus} were also found to switch from nocturnal to diurnal behaviour in response to fox activity (Fenn \& Macdonald 1995). For cats in our study, a switch to diurnal behaviour where fox activity was high in the dry vegetation types may have been facilitated by the higher abundance of reptiles in these habitat types relative to wet forests (which are mostly diurnal; Woinarski \emph{et al.} 2018). At wet forest sites with high fox activity, cats concentrated their activity away from sunrise and sunset towards midnight, despite a diurnal shift appearing to similarly reduce the risk of a fox encounter. In this situation, we expect becoming more nocturnal to be favourable over a diurnal shift because this is when small mammals are active and cats would be least visible to foxes {[}cats here mostly had black or grey coats{]}. Overall fox activity was twice as low in the wet forests relative to dry habitat types, and so cats were likely under less pressure to radically alter their diel activity patterns. Understanding how these potential avoidance behaviours impacts native prey is a key research priority to improve invasive predator management.

Cats may have avoided foxes in time, but we saw no sign of spatial avoidance (i.e.~no fine-scale negative association between fox and cat overall activity (Fig. \ref{fig:diel-cat-fox}b:c). Overlap in spatial activity at fine scales (i.e., at the camera-trap sites) may have been facilitated by temporal avoidance (Kronfeld-Schor \& Dayan 2003). However, we considered spatial avoidance across the entire survey duration (averaging 47 days). Cats may indeed avoid foxes spatially, but transiently on the scale of hours to days---after all, how is a cat to know where to avoid a fox without encountering signs of one? Short-term spatial avoidance is quite plausible given foxes mark territories using scats and odours, which cats could tangibly associate with high risk shortly after. Temporary spatial avoidance could be tested using decay curves (e.g., Niedballa \emph{et al.} 2019). Alternatively, no sign of spatial avoidance between these invasive predators could also be an artefact of the quality of camera set-up and hence detectability. The 3667 camera-traps were deployed by numerous people, and the quality of set-ups differed considerably in terms of detecting predators. Camera-traps that angled only slightly downwards, rather than upwards or strongly downwards, seemed most effective at detecting both predator species (M.W Rees, personal observation). Predator interactions are routinely inferred through spatial associations between species; however such analyses are subject to numerous pitfalls which can make inference unreliable (reviewed in Blanchet, Cazelles, \& Gravel 2020).

A distinction of our study from others is that we modelled potential avoidance behaviours in a simple predator guild, where apex predator activity was artificially manipulated. This reduces potential bias from differences in niche preferences and the unmodelled impacts of other predators in the system. We also included replication across different habitat types. However, because our study did not consider associations with prey species, we cannot distinguish whether changes in cat diel activity patterns were the result of direct fox avoidance or indirect associations with shared prey. For example, low fox activity may promote the availability of a preferred shared prey species with a diel pattern which differs from those of cats on average, and as a result, cats might shift diel activity patterns at sites with low fox activity to more closely match those of the more abundant prey species. We would expect introduced European rabbits \emph{Oryctolagus cuniculus} and hares \emph{Lepus europaeus} (which are diurnal) to be particularly likely to induce such a response in dry vegetation types (McGregor \emph{et al.} 2020; Stobo-Wilson \emph{et al.} 2020), however, rabbits and hares are rare within the natural vegetation of the landscapes we surveyed (only ever being detected at 40 of 1232 sites) and there is little evidence that predation by foxes suppresses rabbit populations (Norbury \& Jones 2015; Scroggie \emph{et al.} 2018). While of particular interest, whether temporal fox-cat interactions are direct or indirect does not change the resulting impact on native prey, and hence the outcomes of fox management.

@@

Animal diel activity patterns can be complex, varying across space, habitat types and threat-levels. Despite telling an important story about how animals interact with each other and the environment, detection times are commonly discarded from statistical analyses of camera-trap data. In the rare instances that they are considered, diel activity patterns are predominantly estimated at the population-level, overlooking finer-scale behaviours that can affect fitness, survival and ecosystem-impacts. Our results demonstrate the importance of (a) considering diel activity in regards to species interactions, (b) modelling \emph{changes} in animal behaviour rather than overlap with other species, and (c) testing avoidance behaviours within a joint spatiotemporal framework. Our study adds to the limited body of evidence that top predators can produce a landscape of fear which is powerful enough to reverse the diel activity patterns of subordinate species (Kronfeld-Schor \& Dayan 2003).

@@

\leavevmode\hypertarget{ref-brashares2010ecological}{}%
Brashares, J.S., Prugh, L.R., Stoner, C.J. \& Epps, C.W. (2010) Ecological and conservation implications of mesopredator release. \emph{Trophic cascades: predators, prey, and the changing dynamics of nature}, 221--240.

@@

\leavevmode\hypertarget{ref-daan1981adaptive}{}%
Daan, S. (1981) Adaptive daily strategies in behavior. \emph{Biological rhythms} (ed J. Aschoff), pp. 275--298. Springer US, Boston, MA.

@@

@@

\leavevmode\hypertarget{ref-kronfeld2003partitioning}{}%
Kronfeld-Schor, N. \& Dayan, T. (2003) Partitioning of time as an ecological resource. \emph{Annual review of ecology, evolution, and systematics}, \textbf{34}, 153--181.

@@

\leavevmode\hypertarget{ref-lima1990behavioral}{}%
Lima, S.L. \& Dill, L.M. (1990) Behavioral decisions made under the risk of predation: A review and prospectus. \emph{Canadian journal of zoology}, \textbf{68}, 619--640.

@@

\leavevmode\hypertarget{ref-miller2014finite}{}%
Miller, D.L. \& Wood, S.N. (2014) Finite area smoothing with generalized distance splines. \emph{Environmental and ecological statistics}, \textbf{21}, 715--731.

@@

\leavevmode\hypertarget{ref-norbury2015pests}{}%
Norbury, G. \& Jones, C. (2015) Pests controlling pests: Does predator control lead to greater european rabbit abundance in australasia? \emph{Mammal Review}, \textbf{45}, 79--87.

@@

\leavevmode\hypertarget{ref-woinarski2018reptiles}{}%
Woinarski, J., Murphy, B., Palmer, R., Legge, S., Dickman, C., Doherty, T., Edwards, G., Nankivell, A., Read, J. \& Stokeld, D. (2018) How many reptiles are killed by cats in Australia? \emph{Wildlife Research}, \textbf{45}, 247--266.

>>>>

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1.1in]{geometry}
\bibliographystyle{elsarticle-harv}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Spatial variation in predator diel activity patterns -- feral cats avoid red foxes in time, not space},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

@@

\linenumbers

@@

We compiled data from multiple smaller-scale camera-trap studies across two regions in south-west Victoria, Australia: the Glenelg region and Otway Ranges (Fig. \ref{fig:diel-map}). Introduced foxes and cats are the only medium-large functional mammalian terrestrial carnivores here: native dingoes \emph{Canis familiaris} are long-absent throughout, while tiger quolls \emph{Dasyurus maculatus} are long-absent in the Glenelg region and likely functionally extinct in the Otway Ranges (last confirmed sighting in 2014). In broad sections of each region, government land managers conduct ongoing targeted lethal fox control for biodiversity conservation. Poison-baits containing 3 mg of sodium fluroacetate (`1080') are buried at a depth of 12 - 15 cm at 1-km intervals along accessible forest tracks and roads. Different road densities result in variable densities of poison-baits. Managers also frequently implement prescribed fire across both regions, primarily to reduce fuel loads to prevent large wildfires.

@@

In the Glenelg region, large patches of natural vegetation are fragmented by pastoral farming and residential properties (Fig. \ref{fig:diel-map}). Foxes in three distinct forest blocks in this region have been subject to poison-baiting since October 2005, with fortnightly bait replacements (Robley \emph{et al.} 2014). These forest blocks, along with three similar, unbaited forest blocks to the north have been simultaneously surveyed annually under the `Glenelg Ark' fox control program since 2005 (40 sites per block; Robley \emph{et al.} 2020). Hair-tubes were used to monitor species from 2005 - 2013 (presented in Robley \emph{et al.} 2014), replaced by camera-traps from 2013; here we present camera-trap data from 2013 - 2019 (Robley \emph{et al.} 2020). We also included a further 425 camera-trap deployments at unique locations from early 2018 (M.W.R. PhD surveys). This totals 2039 camera-trap deployments in the Glenelg region, collected in a control-impact experimental design (foxes had been continuously controlled for at 8 - 14 years in the treatment landscapes at the time of these surveys).

@@

\hypertarget{camera-trap-set-ups}{%
\subsubsection{Camera-trap set-ups}\label{camera-trap-set-ups}}

@@

\hypertarget{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-vegetation-types-model-2}{%
\subsubsection{How does predator overall activity and diel activity patterns vary across vegetation types? (model 2)}\label{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-vegetation-types-model-2}}

@@

\hypertarget{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-space-model-1-1}{%
\subsubsection{How does predator overall activity and diel activity patterns vary across space? (model 1)}\label{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-space-model-1-1}}

@@

Variation in diel activity patterns across space, as well as differences in overall activity between the predators, was strongest in the Otway Ranges. For example, overall fox activity (Fig. \ref{fig:diel-space-o-fox}) and diel activity pattern strength (Fig. \ref{fig:diel-space}b) were lowest in the south-west Otway Ranges, while feral cat overall activity (Fig. \ref{fig:diel-space-o-cat}) and diel activity pattern strength (Fig. \ref{fig:diel-space}b) were highest in that subregion.

\hypertarget{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-vegetation-types-model-2-1}{%
\subsubsection{How does predator overall activity and diel activity patterns vary across vegetation types? (model 2)}\label{how-does-predator-overall-activity-and-diel-activity-patterns-vary-across-vegetation-types-model-2-1}}

@@

Diel activity patterns for foxes were similar across all vegetation type except wet forests; in wet forests, foxes were consistently active throughout the daily cycle (Fig. \ref{fig:diel-veg}a). On the other hand, cats were nocturnal (and most active) in wet forests, but largely crepuscular in all other vegetation types (Fig. \ref{fig:diel-veg}b). For both predators, the random effect for region (Glenelg or Otways) in the vegetation models shrank to near-zero, indicating all variation between the regions was explained by the vegetation covariate and site random intercept.

\hypertarget{do-feral-cats-avoid-foxes-in-space-or-time-model-3-1}{%
\subsubsection{Do feral cats avoid foxes in space or time? (model 3)}\label{do-feral-cats-avoid-foxes-in-space-or-time-model-3-1}}

@@

\caption{Variation in mean feral cat \textit{Felis catus} activity (a) and associated uncertainty estimates (b) in response to count of ’independent’ red fox \textit{Vulpes vulpes} detections (log-transformed and survey effort adjusted) across each ’habitat type’ in south-west Victoria, Australia (model 3). Grey vertical lines represent average sunrise and sunset times. In the Glenelg region, there were more feral cat detections where there were more fox detections, but cat peak diel activity shifted from crepuscular night to pre-dawn and midday (a). In the Otway Ranges, feral cat activity also peaked during the day where fox activity was high in dry vegetation types (b), but was more nocturnal where fox activity was high in the rainforests and wet forests (c).}\label{fig:diel-cat-fox}
\end{figure}

@@

A key question in ecological theory is whether animals are evolutionary hardwired to occupy particular temporal niches, or have circadian rhythms that are responsive to changing environmental conditions and interactions with other species (Schoener 1974; Daan 1981; Lima \& Dill 1990). Here we demonstrate that diel activity patterns are not fixed, but vary across space based on landscape context and fear. In our study, sympatric invasive predators had similar diel activity patterns when averaged across broad regions (i.e., high circular overlap; Fig. \ref{fig:diel-space-time-marginal}; as did Roshier \& Carter 2021), but behaviours varied considerably within landscapes. Fox daily activity patterns were most strongly tied to the daily cycle in dry habitat types, but showed little diel activity pattern in the wet forests (Fig. \ref{fig:diel-veg}a). In contrast, cats were mostly nocturnal in wet forests but crepuscular in dry vegetation types (Fig. \ref{fig:diel-veg}b). Within broad habitat types, cats altered their diel activity patterns at sites with higher fox activity (Fig. \ref{fig:diel-cat-fox}). Control programs that reduce invasive fox activity are therefore likely to change cat diel activity patterns, which in turn may alter cat impacts on native prey species. Quantifying changes in diel activity patterns provides important context for understanding species interactions, which is key for effective ecosystem management (Gaynor \emph{et al.} 2021).

Shifting diel activity patterns may facilitate spatial coexistence of dominant and subordinate species (Carothers \& Jaksić 1984). For cats, altering diel activity patterns to less-preferred times of day may be worthwhile to persist in high-quality habitat. Few studies have demonstrated predator-induced shifts in diel activity (Kronfeld-Schor \& Dayan 2003), but notably, ship rats \emph{Rattus norvegicus} were also found to switch from nocturnal to diurnal behaviour in response to fox activity (Fenn \& Macdonald 1995) and a similar nocturnal-diurnal shift was observed in American mink \emph{Neovison vison} following the recolonisation of native predators (Harrington \emph{et al.} 2009). For cats in our study, a switch to diurnal behaviour where fox activity was high in the dry vegetation types may have been facilitated by the higher abundance of reptiles in these habitat types relative to wet forests (which are mostly diurnal; Woinarski \emph{et al.} 2018). At wet forest sites with high fox activity, cats concentrated their activity away from sunrise and sunset towards midnight, despite a diurnal shift appearing to similarly reduce the risk of a fox encounter. In this situation, we expect becoming more nocturnal to be favourable over a diurnal shift because this is when small mammals are active and cats would be least visible to foxes (cats here mostly had black or grey coats). Overall fox activity was twice as low in the wet forests relative to dry habitat types, and so cats were likely under less pressure to radically alter their diel activity patterns. Understanding how these potential avoidance behaviours impacts native prey is a key research priority to improve invasive predator management.

Cats may have avoided foxes in time, but we saw no sign of spatial avoidance (i.e.~no fine-scale negative association between fox and cat overall activity (Fig. \ref{fig:diel-cat-fox}b:c). Overlap in spatial activity at fine scales (i.e., at the level of camera-trap sites) may have been facilitated by temporal avoidance (Kronfeld-Schor \& Dayan 2003). However, we considered spatial avoidance across the entire survey duration (averaging 47 days). Cats may indeed avoid foxes spatially, but transiently on the scale of hours to days---after all, how is a cat to know where to avoid a fox without encountering signs of one? Short-term spatial avoidance is quite plausible given foxes mark territories using scats and odours, which cats could tangibly associate with high risk shortly after. Temporary spatial avoidance could be tested using decay curves (e.g., Niedballa \emph{et al.} 2019). Alternatively, no sign of spatial avoidance between these invasive predators could also be an artefact of the quality of camera set-up and hence detectability. The 3667 camera-traps were deployed by numerous people, and the quality of set-ups differed considerably in terms of detecting predators. Camera-traps that angled only slightly downwards, rather than upwards or strongly downwards, seemed most effective at detecting both predator species (M.W Rees, personal observation). Predator interactions are routinely inferred through spatial associations between species; however such analyses are subject to numerous pitfalls which can make inference unreliable (reviewed in Blanchet, Cazelles, \& Gravel 2020).\\
A distinction of our study from others is that we modelled potential avoidance behaviours in a simple predator guild, where apex predator activity was artificially manipulated. This reduces potential bias from differences in niche preferences and the unmodelled impacts of other predators in the system. We also included replication across different habitat types. However, because our study did not consider associations with prey species, we cannot distinguish whether changes in cat diel activity patterns were the result of direct fox avoidance or indirect associations with shared prey. For example, low fox activity may promote the availability of a preferred shared prey species with a diel pattern which differs from those of cats on average, and as a result, cats might shift diel activity patterns at sites with low fox activity to more closely match those of the more abundant prey species. We would expect introduced European rabbits \emph{Oryctolagus cuniculus} and hares \emph{Lepus europaeus} (which are diurnal) to be particularly likely to induce such a response in dry vegetation types (McGregor \emph{et al.} 2020; Stobo-Wilson \emph{et al.} 2020), however, rabbits and hares are rare within the natural vegetation of the landscapes we surveyed (only ever being detected at 40 of 1232 sites) and there is little evidence that predation by foxes suppresses rabbit populations (Norbury \& Jones 2015; Scroggie \emph{et al.} 2018). While of particular interest, whether temporal fox-cat interactions are direct or indirect does not change the resulting impact on native prey, and hence the outcomes of fox management.

@@

Animal diel activity patterns can be complex, varying across space, habitat types and threat-levels. Despite telling an important story about how animals interact with each other and the environment, detection times are commonly discarded from statistical analyses of camera-trap data. In the rare instances that they are considered, diel activity patterns are predominantly estimated at the population-level, overlooking finer-scale behaviours that can affect fitness, survival and ecosystem-impacts. Our results demonstrate the importance of (a) considering diel activity in regards to species interactions, (b) modelling changes in animal behaviour rather than overlap with other species, and (c) testing avoidance behaviours within a joint spatiotemporal framework. Our study adds to the limited body of evidence that top predators can produce a landscape of fear which is powerful enough to reverse the diel activity patterns of subordinate species (Kronfeld-Schor \& Dayan 2003).

@@

\leavevmode\hypertarget{ref-brashares2010ecological}{}%
Brashares, J.S., Prugh, L.R., Stoner, C.J. \& Epps, C.W. (2010) Ecological and conservation implications of mesopredator release. \emph{Trophic cascades: Predators, prey, and the changing dynamics of nature} pp. 221--240. Island Press.

@@

\leavevmode\hypertarget{ref-daan1981adaptive}{}%
Daan, S. (1981) Adaptive daily strategies in behavior. \emph{Biological rhythms} pp. 275--298. Springer Publishing.

@@

\leavevmode\hypertarget{ref-harrington2009impact}{}%
Harrington, L.A., Harrington, A.L., Yamaguchi, N., Thom, M.D., Ferreras, P., Windham, T.R. \& Macdonald, D.W. (2009) The impact of native competitors on an alien invasive: Temporal niche shifts to avoid interspecific aggression. \emph{Ecology}, \textbf{90}, 1207--1216.

@@

\leavevmode\hypertarget{ref-kronfeld2003partitioning}{}%
Kronfeld-Schor, N. \& Dayan, T. (2003) Partitioning of time as an ecological resource. \emph{Annual Review of Ecology, Evolution, and Systematics}, \textbf{34}, 153--181.

@@

\leavevmode\hypertarget{ref-lima1990behavioral}{}%
Lima, S.L. \& Dill, L.M. (1990) Behavioral decisions made under the risk of predation: A review and prospectus. \emph{Canadian Journal of Zoology}, \textbf{68}, 619--640.

@@

\leavevmode\hypertarget{ref-miller2014finite}{}%
Miller, D.L. \& Wood, S.N. (2014) Finite area smoothing with generalized distance splines. \emph{Environmental and Ecological Statistics}, \textbf{21}, 715--731.

@@

\leavevmode\hypertarget{ref-norbury2015pests}{}%
Norbury, G. \& Jones, C. (2015) Pests controlling pests: Does predator control lead to greater European rabbit abundance in Australasia? \emph{Mammal Review}, \textbf{45}, 79--87.

@@

\leavevmode\hypertarget{ref-woinarski2018reptiles}{}%
Woinarski, J.C.Z., Murphy, B., Palmer, R., Legge, S., Dickman, C., Doherty, T., Edwards, G., Nankivell, A., Read, J. \& Stokeld, D. (2018) How many reptiles are killed by cats in Australia? \emph{Wildlife Research}, \textbf{45}, 247--266.
